---
title: "MKR-ETHB-Visualizations"
author: "Charliemarketplace"
date: "`r Sys.Date()`"
output:
  html_document:
    css: "styles.css"
    includes:
      in_header: header.html
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


# Intro 

This report details the exploratory data analysis of Key ETH-B Collateral Type 
Vault interactions by users of the Maker Protocol used to inspire & review the 
final MKR ETH-B Elasticity report.

```{r, message=FALSE, warning=FALSE}
library(reactable)
library(plotly)
library(dplyr)
library(lubridate)
```

# Data 

Key data includes: 

- known parameter changes to the ETH-B vault: 
    - JUG.ilks.duty: vault stability fee
    - DOG.ilks.chop: liquidation fee
    - DOG.ilks.hole: liquidation parameter
    - VAT.ilks.dust:	Minimum vault size for # of DAI able to be minted
- Vault Interactions
    - Vaults created with ETH-B as the Collateral type (`ilk`)
    - ETH Deposits to ETH-B Vaults
    - DAI Mints (borrows) via ETH-B Vaults
    - DAI Repayments to ETH-B Vaults
    - ETH Withdrawals from ETH-B Vaults
    - DAI/ETH Liquidations of ETH-B Vaults
- Macro Factors
    - Changes to ETH's Price, specifically its Relative Strength Index (RSI)
    - Demand for ETH's blockspace, i.e., gas (in gwei).


```{r}
parameter_changes <- read.csv("ETHB-Parameter-Changes.csv", row.names = NULL)

ethb_vault_creations <- read.csv("ethb_vault_creations.csv", 
                                 colClasses = "character", row.names = NULL)
ethb_vault_deposits <- read.csv("ethb_vault_deposits.csv", 
                                colClasses = "character", row.names = NULL)
ethb_dai_borrows <- read.csv("ethb_dai_borrows.csv", 
                             colClasses = "character", row.names = NULL)
ethb_dai_repayments <- read.csv("ethb_dai_repayments.csv",
                                colClasses = "character", row.names = NULL)
ethb_vault_withdrawals <- read.csv("ethb_vault_withdrawals.csv",
                                   colClasses = "character", row.names = NULL)
ethb_vault_liquidations <- read.csv("ethb_vault_liquidations.csv",
                                    colClasses = "character", row.names = NULL)

```

# ETH-B History

```{r}
fee_changes <- parameter_changes[parameter_changes$PARAMETER == 'JUG.ilks.duty', c("BLOCK","TIMESTAMP", "FROM_VALUE","TO_VALUE")]

fee_changes$direction <- factor(fee_changes$TO_VALUE > fee_changes$FROM_VALUE, 
                                levels = c(TRUE, FALSE),
                                labels = c("UP","DOWN"))


```

ETH-B was initiated on `r fee_changes$TIMESTAMP[1]` with a stability fee of `r fee_changes$TO_VALUE[1]`.

There have been a total of `r nrow(fee_changes) - 1` fee changes from the initial fee as of the 
analysis cutoff at Ethereum mainnet block 16,400,000 (2023-01-13).

A total of `r nrow(ethb_vault_creations)` vaults were created over this time period.

## Stability Fee Overlay

```{r}
reactable(fee_changes, defaultPageSize = 15)

# for better plotting

fee_changes$TIMESTAMP_posixct <- as.POSIXct(fee_changes$TIMESTAMP, tz = "UTC")
fee_changes$TIMESTAMP_date <- as.Date(fee_changes$TIMESTAMP, tz = "UTC")


# for better plotting
parameter_changes$TIMESTAMP_posixct <- as.POSIXct(parameter_changes$TIMESTAMP, tz = "UTC")
parameter_changes$TIMESTAMP_date <- as.Date(parameter_changes$TIMESTAMP, tz = "UTC")

debt_changes <- parameter_changes[parameter_changes$PARAMETER %in% 
                                    c('DC-IAM.ilks.line','VAT.ilks.line'), ]


plot_ly() %>% add_trace(data = fee_changes,
                        x = ~TIMESTAMP_date, 
                        y = ~TO_VALUE, 
            line = list(shape = 'hv'), # stepwise changes
         text = paste('Block: ', fee_changes$BLOCK,
                        '<br>Timestamp: ', fee_changes$TIMESTAMP_posixct,
                        '<br>From: ',fee_changes$FROM_VALUE,
                        'To: ', fee_changes$TO_VALUE),
         hoverinfo = 'text',
        type = "scatter", mode = "markers+lines") %>% 
  layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "Stability Fee"), 
    title = list(text = "Historical Changes to ETH-B Stability Fee", y = 0.975)
  )


```

## Creations

```{r}
daily_vaults <- ethb_vault_creations[, c("BLOCK_NUMBER", "BLOCK_TIMESTAMP", "VAULT_NUMBER")]
daily_vaults$day <- as.Date(daily_vaults$BLOCK_TIMESTAMP, tz = "UTC")
daily_vaults <- daily_vaults %>% group_by(day) %>% summarise(num_vaults = n())

plot_ly() %>% add_trace(data = daily_vaults,
                        x = ~day, 
                        y = ~num_vaults, 
         text = paste('Date: ', daily_vaults$day,
                        '<br># ETH-B Vaults:', daily_vaults$num_vaults),
         hoverinfo = 'text',
        type = "scatter", mode = "markers+lines", name = 'New Vaults') %>% 
  layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# Vaults"), 
    title = list(text = "Daily ETH-B Vaults Created", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```


## Deposits

```{r}
daily_deposits <- ethb_vault_deposits[ , c("BLOCK_NUMBER","BLOCK_TIMESTAMP", "DEPOSIT_AMOUNT_ADJ")]

daily_deposits$day <- as.Date(daily_deposits$BLOCK_TIMESTAMP, tz = 'UTC')

daily_deposits <- daily_deposits %>% group_by(day) %>% 
  summarise(num_deposits = n(), 
            sum_eth = sum(as.numeric(DEPOSIT_AMOUNT_ADJ)))


monthly_deposits <- daily_deposits %>% 
  mutate(yearmonth = substr(day, 0, 7)) %>% 
  mutate(yearmonth = factor(yearmonth, 
                            levels = unique(yearmonth))) %>% 
  group_by(yearmonth) %>% 
   summarise(num_deposits = sum(num_deposits), 
            sum_eth = sum(as.numeric(sum_eth))) 

```

### Daily: # Deposits 

```{r}
plot_ly() %>% add_trace(
  data = daily_deposits, 
  x = ~day, 
  y = ~num_deposits,
  size = ~sum_eth,
  type = "scatter",
  mode = "markers", 
  name = "Sized by ETH Amount") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# of Deposits"), 
    title = list(text = "Daily ETH-B Vaults Deposits", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```

### Daily: # ETH  

```{r}
plot_ly() %>% add_trace(
  data = daily_deposits, 
  x = ~day, 
  y = ~sum_eth,
  size = ~num_deposits,
  type = "scatter",
  mode = "markers", 
  name = "Sized by # Deposits") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# ETH Deposited"), 
    title = list(text = "Daily ETH-B Vaults ETH Deposited", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100000,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))
  
```


### Monthly: # Deposits 

```{r}
plot_ly() %>% add_trace(
  data = monthly_deposits, 
  x = ~yearmonth, 
  y = ~num_deposits,
  color = ~sum_eth,
  textposition = "none",
  text = paste('Date: ', monthly_deposits$yearmonth,
               '<br># Deposits: ',monthly_deposits$num_deposits,
               '<br>ETH Deposited: ', format(floor(monthly_deposits$sum_eth), big.mark = ",")),
  hoverinfo = 'text',
  type = "bar",
  name = "# Deposits") %>% 
  add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*1000,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))  %>% 
  layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# Deposits"), 
    title = list(text = "Monthly ETH-B Vaults # Deposits", y = 0.975)) %>% 
  colorbar(title = list(text = "# ETH"))


```

### Monthly: # ETH  

```{r, warning=FALSE, message=FALSE}
plot_ly() %>% add_trace(
  data = monthly_deposits, 
  x = ~yearmonth, 
  y = ~sum_eth,
  color = ~num_deposits,
  type = "bar", 
   textposition = "none",
  text = paste('Date: ', monthly_deposits$yearmonth,
               '<br># Deposits: ',monthly_deposits$num_deposits,
               '<br>ETH Deposited: ', format(floor(monthly_deposits$sum_eth), big.mark = ",")),
  hoverinfo = 'text',
  name = "ETH Deposited") %>% 
   layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# ETH Deposited"), 
    title = list(text = "Monthly ETH-B Vaults ETH Deposited", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100000,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv')) %>% 
    colorbar(title = list(text = "# Deposits"))

```

## Mints

```{r}
daily_mints <- ethb_dai_borrows[, c("BLOCK_NUMBER","BLOCK_TIMESTAMP", "DAI_MINTED")]

daily_mints$day <- as.Date(daily_mints$BLOCK_TIMESTAMP, tz = 'UTC')

daily_mints <- daily_mints %>% group_by(day) %>% 
  summarise(num_mints = n(), 
            sum_dai = sum(as.numeric(DAI_MINTED)))

monthly_mints <- daily_mints %>% 
  mutate(yearmonth = substr(day, 0, 7)) %>% 
  mutate(yearmonth = factor(yearmonth, 
                            levels = unique(yearmonth))) %>% 
  group_by(yearmonth) %>% 
   summarise(num_mints = sum(num_mints), 
            sum_dai = sum(as.numeric(sum_dai))) 


```

## Daily: # Mints

```{r}

plot_ly() %>% add_trace(
  data = daily_mints, 
  x = ~day, 
  y = ~num_mints,
  size = ~sum_dai,
  type = "scatter",
  mode = "markers", 
  name = "Sized by DAI Amount") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# of Mints"), 
    title = list(text = "Daily ETH-B Vault # of Mints", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```

## Daily: # DAI

```{r}

plot_ly() %>% add_trace(
  data = daily_mints, 
  x = ~day, 
  y = ~sum_dai,
  size = ~num_mints,
  type = "scatter",
  mode = "markers", 
  name = "Sized by # Mints") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "Dai Minted"), 
    title = list(text = "Daily ETH-B Vault Amount Dai Minted", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100*1e6,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```

### Monthly: # Mints 

```{r}
plot_ly() %>% add_trace(
  data = monthly_mints, 
  x = ~yearmonth, 
  y = ~num_mints,
  color = ~sum_dai,
  textposition = "none",
  text = paste('Date: ', monthly_mints$yearmonth,
               '<br># Mints: ',monthly_mints$num_mints,
               '<br>Dai Minted: ', format(floor(monthly_mints$sum_dai), big.mark = ",")),
  hoverinfo = 'text',
  type = "bar",
  name = "# Mints") %>% 
  add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*1000*10,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))  %>% 
  layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# Mints"), 
    title = list(text = "Monthly ETH-B Vaults # Mints", y = 0.975)) %>% 
  colorbar(title = list(text = "# DAI"))


```

### Monthly: # DAI  

```{r, warning=FALSE, message=FALSE}
plot_ly() %>% add_trace(
  data = monthly_mints, 
  x = ~yearmonth, 
  y = ~sum_dai,
  color = ~num_mints,
  type = "bar", 
   textposition = "none",
  text = paste('Date: ', monthly_mints$yearmonth,
               '<br># mints: ',monthly_mints$num_mints,
               '<br>DAI Minted: ', format(floor(monthly_mints$sum_dai), big.mark = ",")),
  hoverinfo = 'text',
  name = "DAI Minted") %>% 
   layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# DAI Minted"), 
    title = list(text = "Monthly ETH-B Vaults DAI Minted", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*1000e6,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv')) %>% 
    colorbar(title = list(text = "# Mints"))

```

## Repayments 

```{r}
daily_repays <- ethb_dai_repayments[, c("BLOCK_NUMBER","BLOCK_TIMESTAMP", "DAI_REPAYED")]

daily_repays$day <- as.Date(daily_repays$BLOCK_TIMESTAMP, tz = 'UTC')

daily_repays <- daily_repays %>% group_by(day) %>% 
  summarise(num_repays = n(), 
            sum_dai = sum(as.numeric(DAI_REPAYED)))

monthly_repays <- daily_repays %>% 
  mutate(yearmonth = substr(day, 0, 7)) %>% 
  mutate(yearmonth = factor(yearmonth, 
                            levels = unique(yearmonth))) %>% 
  group_by(yearmonth) %>% 
   summarise(num_repays = sum(num_repays), 
            sum_dai = sum(as.numeric(sum_dai))) 


```

## Daily: # Repays

```{r}

plot_ly() %>% add_trace(
  data = daily_repays, 
  x = ~day, 
  y = ~num_repays,
  size = ~sum_dai,
  type = "scatter",
  mode = "markers", 
  name = "Sized by DAI Amount") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# of Repays"), 
    title = list(text = "Daily ETH-B Vault # of Repays", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```

## Daily: # DAI

```{r}

plot_ly() %>% add_trace(
  data = daily_repays, 
  x = ~day, 
  y = ~sum_dai,
  size = ~num_repays,
  type = "scatter",
  mode = "markers", 
  name = "Sized by # Repays") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "Dai Repayed"), 
    title = list(text = "Daily ETH-B Vault Amount Dai Repaid", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100*1e6,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```


### Monthly: # Repays 

```{r}
plot_ly() %>% add_trace(
  data = monthly_repays, 
  x = ~yearmonth, 
  y = ~num_repays,
  color = ~sum_dai,
  textposition = "none",
  text = paste('Date: ', monthly_repays$yearmonth,
               '<br># repays: ',monthly_repays$num_repays,
               '<br>Dai Repaid: ', format(floor(monthly_repays$sum_dai), big.mark = ",")),
  hoverinfo = 'text',
  type = "bar",
  name = "# Repays") %>% 
  add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*1000*10,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))  %>% 
  layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# Repays"), 
    title = list(text = "Monthly ETH-B Vaults # Repays", y = 0.975)) %>% 
  colorbar(title = list(text = "# DAI"))


```

### Monthly: # DAI  

```{r, warning=FALSE, message=FALSE}
plot_ly() %>% add_trace(
  data = monthly_repays, 
  x = ~yearmonth, 
  y = ~sum_dai,
  color = ~num_repays,
  type = "bar", 
   textposition = "none",
  text = paste('Date: ', monthly_repays$yearmonth,
               '<br># Repays: ',monthly_repays$num_repays,
               '<br>DAI Repaid: ', format(floor(monthly_repays$sum_dai), big.mark = ",")),
  hoverinfo = 'text',
  name = "DAI Repaid") %>% 
   layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# DAI Repaid"), 
    title = list(text = "Monthly ETH-B Vaults DAI Repaid", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*1000e6,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv')) %>% 
    colorbar(title = list(text = "# Repays"))

```


## Withdrawals 


```{r}
daily_withdraws <- ethb_vault_withdrawals[ , c("BLOCK_NUMBER","BLOCK_TIMESTAMP", "WITHDRAWN_AMOUNT_ADJ")]

daily_withdraws$day <- as.Date(daily_withdraws$BLOCK_TIMESTAMP, tz = 'UTC')

daily_withdraws <- daily_withdraws %>% group_by(day) %>% 
  summarise(num_withdraws = n(), 
            sum_eth = sum(as.numeric(WITHDRAWN_AMOUNT_ADJ)))


monthly_withdraws <- daily_withdraws %>% 
  mutate(yearmonth = substr(day, 0, 7)) %>% 
  mutate(yearmonth = factor(yearmonth, 
                            levels = unique(yearmonth))) %>% 
  group_by(yearmonth) %>% 
   summarise(num_withdraws = sum(num_withdraws), 
            sum_eth = sum(as.numeric(sum_eth))) 

```


### Daily: # Withdrawals 

```{r}
plot_ly() %>% add_trace(
  data = daily_withdraws, 
  x = ~day, 
  y = ~num_withdraws,
  size = ~sum_eth,
  type = "scatter",
  mode = "markers", 
  name = "Sized by ETH Amount") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# of Withdrawals"), 
    title = list(text = "Daily ETH-B Vaults Withdrawals", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```

### Daily: # ETH  

```{r}
plot_ly() %>% add_trace(
  data = daily_withdraws, 
  x = ~day, 
  y = ~sum_eth,
  size = ~num_withdraws,
  type = "scatter",
  mode = "markers", 
  name = "Sized by # Withdrawals") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# ETH Withdrawn"), 
    title = list(text = "Daily ETH-B Vaults ETH Withdrawn", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100000,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))
  
```


### Monthly: # Withdrawals 

```{r}
plot_ly() %>% add_trace(
  data = monthly_withdraws, 
  x = ~yearmonth, 
  y = ~num_withdraws,
  color = ~sum_eth,
  textposition = "none",
  text = paste('Date: ', monthly_withdraws$yearmonth,
               '<br># withdraws: ',monthly_withdraws$num_withdraws,
               '<br>ETH Withdrawn: ', format(floor(monthly_withdraws$sum_eth), big.mark = ",")),
  hoverinfo = 'text',
  type = "bar",
  name = "# Withdraws") %>% 
  add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*10000,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee %', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))  %>% 
  layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# Withdraws"), 
    title = list(text = "Monthly ETH-B Vaults # Withdraws", y = 0.975)) %>% 
  colorbar(title = list(text = "# ETH"))


```

### Monthly: # ETH  

```{r, warning=FALSE, message=FALSE}
plot_ly() %>% add_trace(
  data = monthly_withdraws, 
  x = ~yearmonth, 
  y = ~sum_eth,
  color = ~num_withdraws,
  type = "bar", 
   textposition = "none",
  text = paste('Date: ', monthly_withdraws$yearmonth,
               '<br># Withdraws: ',monthly_withdraws$num_withdraws,
               '<br>ETH Withdrawn: ', format(floor(monthly_withdraws$sum_eth), big.mark = ",")),
  hoverinfo = 'text',
  name = "ETH Withdrawn") %>% 
   layout(
    xaxis = list(title = "Month", range = c("2020-09-01",'2023-02-01')),
    yaxis = list(title = "# ETH Withdrawals"), 
    title = list(text = "Monthly ETH-B Vaults ETH Withdrawn", y = 0.975)) %>% 
    add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*1000000,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv')) %>% 
    colorbar(title = list(text = "# Withdrawals"))

```

## Liquidations 

```{r}
daily_liqs <- ethb_vault_liquidations[ , 
                                       c("BLOCK_TIMESTAMP",
                                         "VAULT_NUMBER",
                                         "DEBT_TO_COVER_AMOUNT",
                                         "COLLATERAL_LIQUIDATED_AMOUNT",
                                         "DAI_REPAYED_AMOUNT")]

daily_liqs$day <- as.Date(daily_liqs$BLOCK_TIMESTAMP, tz = 'UTC')

daily_liqs <- daily_liqs %>% group_by(day) %>% 
  summarise(num_liqs = n(), 
            sum_eth = sum(as.numeric(COLLATERAL_LIQUIDATED_AMOUNT)),
            sum_debt = sum(as.numeric(DEBT_TO_COVER_AMOUNT)),
            sum_dai = sum(as.numeric(DAI_REPAYED_AMOUNT)),
            )

monthly_liqs <- daily_liqs %>% 
  mutate(yearmonth = substr(day, 0, 7)) %>% 
  mutate(yearmonth = factor(yearmonth, 
                            levels = unique(yearmonth))) %>% 
  group_by(yearmonth) %>% 
   summarise(num_liqs = sum(num_liqs), 
            sum_eth = sum(as.numeric(sum_eth))) 


```

Liquidations will not be thoroughly analyzed as the vast majority of liquidations are of a 
single large vault: # 21348, which had 363 distinct liquidations in chunks. But they 
will be included as ETH withdrawn and DAI debt covered for calculating DAI outstanding.

# Key Metrics 

## DAI oustanding over time vs Debt Limit

- DAI Outstanding: DAI mints - DAI repays - Dai liquidation debt covered
- Debt Ceiling Parameter: `VAT.ilks.line` - automatic increases in allowed debt for vault 
up to the governance imposed Max.
- Debt Max Limit Parameter: `DC-IAM.ilks.line` - governance imposed true maximum debt, 
can be approached slowly automatically by increasing the ceiling.


The key context to note is that from Jan 2021 to June 2021, the ETH-B Ceiling had completely
maxxed out, requiring all mints of DAI to be offset by recent repayments of DAI. Causing a 
significant decrease in activity to ETH-B Vaults as the amount of DAI available to be minted 
was reduced. 

After June 2021, Maker governance dramatically increased the DC-IAM Max Debt Limit 
and since then, automatic increases in the Ceiling (even in high-volatility) had 
plenty of room to allow additional dai minting.

```{r}

plot_ly() %>% add_trace(data = debt_changes,
                        x = ~TIMESTAMP_date, 
                        y = ~TO_VALUE, 
                        color = ~PARAMETER,
            line = list(shape = 'hv'), # stepwise changes
         text = paste('Block: ', debt_changes$BLOCK,
                        '<br>Timestamp: ', debt_changes$TIMESTAMP_posixct,
                        '<br>From: ',debt_changes$FROM_VALUE,
                        'To: ', debt_changes$TO_VALUE),
         hoverinfo = 'text',
        type = "scatter", mode = "markers+lines") %>% 
  layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "Debt Limit"), 
    title = list(text = "Historical Changes to ETH-B Debt Ceiling & Limit", y = 0.975)
  ) %>% 
  add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100*10e6,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```


```{r}
daily_mints$dai_minted <- daily_mints$sum_dai
daily_repays$dai_repaid <- daily_repays$sum_dai

# FULL Join 
mint_repay <- merge(daily_mints[, c("day","dai_minted")], 
      daily_repays[, c("day","dai_repaid")], by = 'day', 
      all.x = TRUE, all.y = TRUE)

dai_outstanding <- merge(mint_repay,  daily_liqs[, c("day", "sum_debt")], 
                         by = "day", all.x = TRUE, all.y = TRUE)


# if nothing happened, impute 0 
dai_outstanding[is.na(dai_outstanding)] <- 0

# Dai Delta = change in dai outstanding attributable to vault
# Dai Outstanding = cumulative sum of deltas 
dai_outstanding <- dai_outstanding %>% 
  mutate(dai_delta = dai_minted - dai_repaid - sum_debt) %>% 
  mutate(dai_outstanding = cumsum(dai_delta))
  
```

The Debt Ceiling, while under the Debt Limit, is flexible to the utilization 
of the ETHB Vault. It is stable in low volatility periods, but spikes in ETHB's utilization, 
(e.g., a large DAI mint) can directly cause or even include in its transaction a change to the ceiling.

```{r}
plot_ly() %>% 
  add_trace(data = dai_outstanding, x = ~day, y = ~dai_outstanding, 
            type = "scatter", mode = "markers+lines", name = "ETHB: Dai Outstanding") %>% 
   layout(
    xaxis = list(title = "Date"),
    yaxis = list(title = "# of DAI"), 
    title = list(text = "Net DAI in circulation from ETHB", y = 0.975)) %>% 
  add_trace(data = debt_changes,
                        x = ~TIMESTAMP_date,
                        y = ~TO_VALUE, 
                        color = ~PARAMETER,
            line = list(shape = 'hv'), # stepwise changes
         text = paste('Block: ', debt_changes$BLOCK,
                        '<br>Timestamp: ', debt_changes$TIMESTAMP_posixct,
                        '<br>From: ',debt_changes$FROM_VALUE,
                        'To: ', debt_changes$TO_VALUE),
         hoverinfo = 'text',
        type = "scatter", mode = "markers+lines") %>% 
  add_trace(x = fee_changes$TIMESTAMP_date, 
            y = ~fee_changes$TO_VALUE*100*10e6,
            text = paste('Date: ', fee_changes$TIMESTAMP_date,
                         '<br>Fee: ', 100*fee_changes$TO_VALUE,"%"),
            hoverinfo = 'text',
            name = 'Stability Fee % (scaled)', 
            type= "scatter",
            mode = "markers+lines", 
            line = list(shape = 'hv'))

```

## Implied Revenue over time

- Implied Revenue = DAI Outstanding * Stability Fee 

```{r}


```


# Macro Factors 

## ETH Price & RSI (Time Level)

## ETH Gas (weighted avg gwei)

# Metrics vs Factors 

## DAI outstanding versus ETH Price

## DAI outstanding versus ETH RSI 

## Implied Revenue versus ETH Price

## Implied Revenue versus ETH RSI

# Next Steps 

In the Elasticity report, a deeper analysis will be done to develop an elasticity
curve for how decisions within Maker's control (i.e., the stability fee) affect 
the desired outcome (Implied Revenue) after adjusting for external factors (i.e., ETH's price & RSI)
to ultimately provide a defined data-driven methodology to support key decision making within MakerDAO.


